---
title: "ATACFlow Report For %s"
author: "Zheng Wei and Wei Zhang"
date: "`r Sys.Date()`"
output: 
    html_document:
        df_print: paged
        toc: true
        toc_float: true
        number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=TRUE, echo=TRUE,message=FALSE}
library(ggplot2)
load("Report.Rdata")

```

# Summary Table

Sequence files below are set as inputs of the pipeline 

```{r cols.print=2, eval=TRUE, echo=FALSE,warning=FALSE}
filelist
```

Summerized infomation on sequence files has been shown showed below. You can see details in later sections

```{r cols.print=2, rows.print=20, eval=TRUE, echo=FALSE,warning=FALSE}
wholesummary
```

# FastQC

# Remove Adapter

```{r cols.print=2, rows.print=20, eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$removeAdapter$getReportVal("adapters")

```

```{r cols.print=2, rows.print=20, eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$removeAdapter$getReportVal("settings")
```

```{r cols.print=2, rows.print=20, eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$removeAdapter$getReportVal("statistics")
```

```{r cols.print=6, rows.print=20, eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$removeAdapter$getReportVal("distribution")
```

# Mapping QC

## Mapping result

```{r eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$bowtie2Mapping$getReportVal("detail")
```

## Library complexity

```{r cols.print=2, eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$libComplexQC$getReportVal("report")
```

## Filted reads component

```{r cols.print=2, eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$sam2Bed$getReportVal("report")
```

## Fraction of reads in peaks (FRiP)

```{r cols.print=2, eval=TRUE, echo=FALSE,warning=FALSE}
atacProcs$fripQC$getReportVal("report")
```

# Fregment size distribution QC



# TSS enrichment QC

# Peak QC

## Blacklist ratio

## DHS ratio



```{r cols.print=3, rows.print=3, eval=TRUE, echo=FALSE}
cars
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
